\textbf{The status of the archaeal and bacterial census: an update}

### 2016-03-16 submission/rebuttal.md

**Line 24: again I think it would be important to explain here why the focus on full length sequences?**

Given the lack of space allowed to us for the Abstract and Importance sections, we have only included this explanation in the Abstract. We also justify this more completely in the main body of the manuscript.

**Line 33: I would argue this is fundamental to more than just ecology**

We would be happy to entertain this argument and possibly add text if the reviewer would provide additional details.

@@

**Line 42. Regarding the deposition of this sequence in 1983 in Genbank at the NCBI.. I don't think this is possible. I think NCBI did not take over Genbank until later. Plus the paper was in 1978 and the precursor toe Genbank existed before 1983 so it would be good to expain the discrepancies.**

We have edited the sentence here and in the Conclusion section to remove emphasis on the dates. Here we point out that it was released with the rest of GenBank in 1983.

**Line 45: "making it the best-represented gene." details and or/ reference needed for this sentence**

We have edited this line to now read "A search of GenBank's nucleotide database reveals that 16S rRNA gene accessions represent nearly one-third of the database..."

@@

**Line 92. I think the only thing that can be said from this is saturation for the types of sites being sampled**

We hope this is implied from the analysis and discussion we have provided in the manuscript.

**Line 104-105. "it is likely that an even larger number of OTUs have yet to be sampled for both domains." Maybe this is a wording thing but it is unclear to me how we can have "started to saturate" yet have an even larger number of OTUs to be sampled**

Since the rarefaction curves are not actually fully saturated at the OTU level (3% threshold), this means there are still OTUs that we expect to observe with new sampling. Consistent with this, the sampled group itself is made up primarily of singletons, meaning that we are only barely detecting most of the OTUs in the first place. Thus, we believe these two sentences do not contradict each other.

@@

**Line 208. "There is concern". Can you clarify - or, perhaps it would be better to say "we are concerned"**

This has been edited as suggested.

**Line 233. Problems with EMIRGE. I think this is possibly true. But not enough evidence is presented in this paper to make this type of statement. What evidence do you have that any errors are coming from EMIRGE?**

It was the high degree of novelty that was observed among the EMIRGE sequences that makes us concerned. Furthermore, considering EMIRGE hasn't been used to sequence a mock community, it is unclear what the error rates are within contigs generated by the approach. We have tried to be conservative here by using “suggests”, and considering this is a combined Results/Discussion, we felt it was appropriate to add some amount of interpretation to the results.

@@

**Line 292 - 322. Much of the text here discusses data from Silva which appears to have mostly come from EMBL. And other text discusses sequences from other sources. This is all fine. But the text as it is written in the main parts of the paper does not provide enough of this for context. I think it would be very beneficial to add a bit more to the main text on the sources of the sequences.**

We feel that this section of text adequately describes the sources of data used in the paper: SILVA, IMG, and the individual EMIRGE datatsets. These are rolled out in the Results section as we describe the main analysis (SILVA) and the newer methods (IMG and EMIRGE).

**Line 337-338. Figshare. Really good to see what is being released in Figshare and Github for this paper. I would further recommend that sequeneces used in this paper also be included in Figshare so that people do not have to rerun the code in order to get access to the sequences (this is what seems to be required)**

We are reluctant to do this since the sequences were not generated by us, and we would like to avoid duplicating an existing public database. In addition, the SILVA sequences are subject to a somewhat odd licensing agreement. The code posted in GitHub describes in detail how to obtain and process the sequences.
